<!--
%\VignetteEngine{knitr::docco_classic}
%\VignetteIndexEntry{rebird vignette}
-->

# Intro to the rebird package

A programmatic interface to the eBird database. Find out more about eBird at [their website](http://ebird.org/content/ebird/).

## Installation

You can install the stable version from CRAN

```{r eval=FALSE}
install.packages("rebird")
```

Or the development version from Github

```{r eval=FALSE}
install.packages("devtools")
library("devtools")
install_github("ropensci/rebird")
```

Then load the package into the R sesssion

```{r}
library("rebird")
```

## Usage

### Sightings at location determined by latitude/longitude

Search for bird occurrences by latitude and longitude point

```{r ebirdgeo1}
out <- ebirdgeo(species = 'spinus tristis', lat = 42, lng = -76)
head(out)
```

Same, but with additional parameter settings, returning only 10 records, including provisional records, and hotspot records. 

```{r ebirdgeo2}
out1 <- ebirdgeo(lat = 42, lng = -76, max = 10, includeProvisional = TRUE, hotspot = TRUE)
head(out1)
nrow(out1)
```


### Recent sightings frm location IDs

Search for bird occurrences for two locations by their IDs 

```{r ebirdloc1}
out2 <- ebirdloc(locID = c('L99381','L99382'))
head(out2)
```

Search by location ID and species name, as well as some additional parameter settings 

```{r ebirdloc2}
ebirdloc(locID = 'L99381', species = 'larus delawarensis', max = 10, provisional = TRUE, hotspot=TRUE)
```


### Recent observations at a region

Search for bird occurrences by region and species name

```{r ebirdregion1}
out3 <- ebirdregion(region = 'US', species = 'Setophaga caerulescens')
head(out3)
```

Search by location ID and species name, as well as some additional parameter settings. Note that we use `US-OH` to represent Ohio within the US. [See possible region values](https://confluence.cornell.edu/display/CLOISAPI/eBird-1.1-RegionCodeReference).

```{r ebirdregion2}
ebirdregion(region = 'US-OH', max = 10, provisional = TRUE, hotspot = TRUE)
```


### Recent observations at hotspots

Search for bird occurrences by region and species name

```{r ebirdhotspot}
ebirdhotspot(locID = c('L99381','L99382'), species = 'larus delawarensis')
```


### Recent notable sightings

Search for notable sightings at a given latitude and longitude

```{r ebirdnotable}
out4 <- ebirdnotable(lat = 42, lng = -70)
head(out4)
```


### eBird taxonomy

Returns a data.frame of all species in the eBird taxonomy for the given parameter inputs

```{r ebirdtax1}
out5 <- ebirdtaxonomy()
head(out5)
```

Search for hybrid species only

```{r ebirdtax2}
out6 <- ebirdtaxonomy(cat="hybrid")
head(out6)
```