<!--
%\VignetteEngine{knitr::docco_classic}
%\VignetteIndexEntry{rebird vignette}
-->

# Intro to the rebird package

A programmatic interface to the eBird database. Find out more about eBird at [their website](http://ebird.org/home/).

## Installation

You can install the stable version from CRAN

```{r eval=FALSE}
install.packages("rebird")
```

Or the development version from Github

```{r eval=FALSE}
install.packages("devtools")
devtools::install_github("ropensci/rebird")
```

Then load the package into the R sesssion

```{r}
library("rebird")
```

## Usage

The [eBird API server](https://documenter.getpostman.com/view/664302/ebird-api-20/2HTbHW) 
has been updated and thus there are a couple major changes in the way `rebird` works.
API requests to eBird now require users to provide an API key, which is linked to your 
eBird user account. 
You can pass it to the 'key' argument in `rebird` functions, but we highly recommend
storing it as an environment variable called EBIRD_KEY in your .Renviron file.
If you don't have a key, you can obtain one from <https://ebird.org/api/keygen>.

You can keep your .Renviron file in your global R home directory (`R.home()`), your user's home
directory (`Sys.getenv("HOME")`), or your current working directory (`getwd()`). Remember
that .Renviron is loaded once when you start R, so if you add your API key to the file you will
have to restart your R session. See <https://csgillespie.github.io/efficientR/r-startup.html> for
more information on R's startup files.

Furthermore, functions now use species codes, rather than scientific names, for species-specific requests.
We've made the switch easy by providing the `species_code` function, which converts a scientific name to
its species code:

```{r speciescode}
species_code('sula variegata')
```

The `species_code` function can be called within other `rebird` functions, or the species code 
can be specified directly.

### Sightings at location determined by latitude/longitude

Search for bird occurrences by latitude and longitude point

```{r ebirdgeo1}
ebirdgeo(species = species_code('spinus tristis'), lat = 42, lng = -76)
```

Same, but with additional parameter settings, returning only 10 records, including provisional records, and hotspot records. 

```{r ebirdgeo2}
ebirdgeo(lat = 42, lng = -76, max = 10, includeProvisional = TRUE, hotspot = TRUE)
```


### Recent sightings from location IDs

Search for bird occurrences for in a hotspot using its ID

```{r ebirdloc1}
ebirdregion(loc = 'L99381')
```

Search by location ID and species name, as well as some additional parameter settings 

```{r ebirdloc2}
ebirdregion(loc = 'L99381', species = species_code('larus delawarensis'), 
         max = 10, provisional = TRUE, hotspot = TRUE)
```


## Nearest observations of a species

Search for a species' occurrences near a given latitude and longitude

```{r nearestobs}
nearestobs(species_code('branta canadensis'), 42, -76)
```


### Recent observations at a region

Search for bird occurrences by region and species name

```{r ebirdregion1}
ebirdregion(loc = 'US', species = species_code('Setophaga caerulescens'))
```

Search by location ID and species name, as well as some additional parameter settings. Note that we use `US-OH` to represent Ohio within the US.

```{r ebirdregion2}
ebirdregion(loc = 'US-OH', max = 10, provisional = TRUE, hotspot = TRUE)
```


### Frequency of observations at hotspots or regions

Obtain historical frequencies of bird occurrences at a given hotspot

```{r ebirdfreq1}
ebirdfreq(loctype = 'hotspots', loc = 'L196159')
```

Same, but in wide format (for making bar charts)

```{r ebirdfreq2}
ebirdfreq(loctype = 'hotspots', loc = 'L196159', long = FALSE)
```

Obtain frequency data for a given state

```{r ebirdfreq3}
ebirdfreq(loctype = 'states', loc = 'CA-BC')
```

Or county

```{r ebirdfreq4}
ebirdfreq(loctype = 'counties', loc = 'CA-BC-GV')
```

Obtain frequency data within a range of years and months

```{r ebirdfreq5}
ebirdfreq(loctype = 'hotspots', loc = 'L196159', startyear = 2010,
          endyear = 2014, startmonth = 1, endmonth = 3)
```


### Recent notable sightings

Search for notable sightings at a given latitude and longitude

```{r ebirdnotable1}
ebirdnotable(lat = 42, lng = -70)
```

or a region

```{r ebirdnotable2}
ebirdnotable(locID = 'US-NY-109')
```

### Historic Observations

Search for historic observations on a date at a region 

```{r ebirdhistoricobs1}
ebirdhistoricobs(loc = 'US-VA-003', date='2019-02-14',max=10)
```

or set of hotspots

```{r ebirdhistoricobs2}
ebirdhistoricobs(loc = 'L196159', date='2019-02-14', fieldSet='full')
```

### eBird taxonomy

Returns a data.frame of all species in the eBird taxonomy for the given parameter inputs

```{r ebirdtax1}
ebirdtaxonomy()
```

Search for hybrid species only

```{r ebirdtax2}
ebirdtaxonomy(cat = "hybrid")
```

## Information on a given region or hotspot

Obtain detailed information on any valid eBird region

```{r ebirdregioninfo1}
ebirdregioninfo("CA-BC-GV")
```

or hotspot

```{r ebirdregioninfo2}
ebirdregioninfo("L196159")
```
