<!--
%\VignetteEngine{knitr::docco_classic}
%\VignetteIndexEntry{rebird vignette}
-->

# Intro to the rebird package

A programmatic interface to the eBird database. Find out more about eBird at [their website](http://ebird.org/content/ebird/).

## Installation

You can install the stable version from CRAN

```{r}
install.packages("rebird")
```

Or the development version from Github

```{r}
install.packages("devtools")
library("devtools")
install_github("ropensci/rebird")
```

Then load the package into the R sesssion

```{r}
library("rebird")
```

## Usage

Below are some code chunks as examples.

```{r hello, results='asis'}
cat('_hello_ **markdown**!', '\n')
```

Normally you do not need any chunk options.

```{r}
1+1
10:1
rnorm(5)^2
strsplit('hello, markdown vignettes', '')
```

Feel free to draw beautiful plots and write math $P(X>x)=\alpha/2$.

```{r}
n=300; set.seed(123)
par(mar=c(4,4,.1,.1))
plot(rnorm(n), rnorm(n), pch=21, cex=5*runif(n), col='white', bg='gray')
```

## How does it work

The **markdown** package (>= v0.6.2) supports custom HTML templates, and the `docco` engine in **knitr** uses a custom template to compile Markdown to HTML:

```{r header, eval=FALSE}
knit2html(..., markdown.HTML.template = system.file('misc', 'docco-template.html', package='knitr'))
```

That is it.